---
title: "The Differential Contact Score Package"
author: "Zachary McCaw"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=1.5*3, fig.height=1.5*2, fig.align="center", fig.path='Figs/',echo=T, warning=F, message=F, cache=T);
# rmarkdown::latex_dependency(name="bm")
```

## Data Format
Experiment is structured as a list of data.frames, one for each chromosome. 
```{r D01, include=T}
# Data
X = DCS::hicExpExample;
# Chromosomes
names(X);
# Data.frame structure
head(X[["chr18"]]);
```

## Border Scores
Calculates a border score by comparing mean signal intensity in the flanking regions with signal intensity in the intervening region. Required arguments are the HiC experiment, names of the columns in each data.frame containing the normalized HiC signal, and the allocation of samples to treatment groups. Output is a list with one data.frame per group. `Border` is the calculated border score. `minPix` and `minProp` are the minimum number and proportion of pixels observed in one of the aggregation regions. 
```{r B01, include=T}
# Data
X = DCS::hicExpExample;
# Samples
s = c("Auxin_rep1","Auxin_rep2","Ctrl_rep1","Ctrl_rep2");
# Sample groupings
g = c(1,1,0,0);
# Calculate border scores
Border = DCS::getBorderScores(X=X,s=s,g=g,cores=2);
# Output structure
plyr::llply(.data=Border,.fun=head);
```

## Visualization

#### Basic heat-map 
Basic HiC heat-map displaying a test statistic, here the log fold change, within each pixel.  
```{r V01, include=T}
library(ggplot2);
# Data
X = DCS::hicPlotExample;
X0 = X[,c("i","j","lfc_1v0")]
X1 = X[,c("i","j","Auxin_log2Mean","Ctrl_log2Mean")];
# Basic HiC plot
Qb = DCS::plotHicBasic(A=X0,f=120,h.u=40,h.d=40,l.u=40,l.d=40);
qb = Qb$Plot;
qb = qb + scale_fill_gradient2(midpoint=0,limits=c(Qb$minSignal,Qb$maxSignal));
print(qb);
```

#### Smoothed Heat-map
HiC heat-map with the signal smoothed against the separation of the interacting regions, i.e. against the distance from the base of the figure.   
```{r V02, include=T}
# Smoothed HiC Plot
Qs = DCS::plotHicSmoothed(A=X,f=120,h.u=40,h.d=40,l.u=40,l.d=40);
qs = Qs$Plot;
qs = qs + scale_fill_gradient2(midpoint=0,limits=c(Qs$minSignal,Qs$maxSignal));
print(qs);
```

#### Grouped Heat-map
HiC heat-maps plotted separately for each treatment group, optionally smoothed.  
```{r V03, include=T}
q = DCS::plotHicGrouped(A=X1,f=120,h.u=40,h.d=40,l.u=40,l.d=40,smooth=T);
print(q);
```