---
title: "Formatting and Class Definitions"
author: "Zachary McCaw"
date: "Updated: 10/30/2017"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Formatting and Class Definitions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r,include=FALSE}
knitr::opts_chunk$set(fig.width=1.5*3, fig.height=1.5*2, fig.align="center", fig.path='Figs/',echo=T, warning=F, message=F, cache=T, results='hold');
# rmarkdown::latex_dependency(name="bm")
```

## Contents

* [Data Formatting](#data-formatting)
* [DCS Objects](#dcs-objects)
* [FS Objects](#fs-objects)

## Data Formatting

#### Extraction from SparseHiC Objects

Input contact matrices for individual chromosomes are expected to have class `dtCMatrix`, defined in the `Matrix` package. If starting from a `sparseHiCdatum`, defined in the `sparseHiC` package, contact matrices of class `dtCMatrix` are available in the `resolutionNamedList`. The `sparseHiC` package is available on  [github](https://github.com/aryeelab/sparseHiC.git). The follow example demonstrates extraction of a contact matrix from a `sparseHiCdatum`.

```{r}
# Installation of sparseHiC package:
# devtools::install_github("aryeelab/sparseHiC");
library(DCS);
# Example of sparse HiC datum
cat("Class of sHicDatum:\n");
class(sHicDatum);
cat("\n");
cat("Available Resolutions:\n");
names(sHicDatum@resolutionNamedList);
cat("\n");
cat("Available Chromosomes at 40kb Resolution:\n");
names(sHicDatum@resolutionNamedList[["40000"]]);
cat("\n");
# Extracting Contact Matrix for Chromosome 1 at 40kb Resolution
M = sHicDatum@resolutionNamedList[["40000"]]$chr1;
cat("Class of Contact Matrix:\n");
class(M);
```

#### Contact Matrix Aggregation

Contact matrices for individual chromosomes are formatted as objects of class `DCSchr`. Chromosomes from the same subject are aggregated into a sample, and samples are combined into an experiment of class `DCSexp`. In the following, a sample of HiC data available from the NCBI GEO, accession **GSE98671**, are considered. `A1` and `A2` contain data from MES cells depleted of CTCF using auxin. `C1` and `C2` contain data from control MES cells. Each is a list of two `dtCMatrix`-class sparse contact matrices for chromosomes `chr1` and `chr2`. The contact matrices for individual chromosomes are formatted into `DCSchr` objects and aggregated into a sample using `createDCSsmp`. The sample level data are combined into an experiment using `createDCSexp`.   

```{r}
# Auxin 
A1 = DCS::Auxin_rep1;
A2 = DCS::Auxin_rep2;
# Control
C1 = DCS::Control_rep1;
C2 = DCS::Control_rep2;
# Combine sparse matrices for individual chromosomes within a sample
Auxin1 = DCS::createDCSsmp(L=A1,chr=c("1","2"),smp="Aux.R1",group=0);
Auxin2 = DCS::createDCSsmp(L=A2,chr=c("1","2"),smp="Aux.R2",group=0);
Ctrl1 = DCS::createDCSsmp(L=C1,chr=c("1","2"),smp="Ctrl.R1",group=1);
Ctrl2 = DCS::createDCSsmp(L=C2,chr=c("1","2"),smp="Ctrl.R2",group=1);
# Combine samples into an experiment
Exp = DCS::createDCSexp(L=list(Auxin1,Auxin2,Ctrl1,Ctrl2));
head(Exp);
```

## Differential Contact Score Objects

#### Structure

An object of class `DCSchr` has the following slots:

* `@chr`, a string naming the chromosome.
* `@Coord`, an $n_{p}$ by 3 matrix of pixel coordinates, where $n_{p}$ is the number of pixels on the chromosome. `@Coord` has columns $(i,j,d)$, where $i$ and $j$ index pixel position, and $d = (j-i)$ is the separation between the interacting loci. Coordinates are in upper-triangular format, i.e. $i \leq j$.
* `@Counts`, an $n_{p}$ by $n_{s}$ matrix, where $n_{s}$ is the number of samples.
* `@groups`, a factor of group assignments for the samples.
* `@samples`, a character vector of sample names.
* `@Stats`, an $n_{p}$ by $n_{q}$ data.frame, where $n_{q}$ is the number of available test statistics.

```{r}
# Extracting chromosome one
A.1 = Exp@Data[["1"]];
# Structure of a DCS chromosome
str(A.1);
```

An object of class `DCSexp` has the following slots:

* `@chrs`, a character vector of chromosome names.
* `@Data`, is a list of `DCSchr`s, with names corresponding to `@chrs`.
* `@groups`, a factor of group assignments for the samples.
* `@samples`, a character vector of sample names.

Note that the `@groups` and `@samples` slots for the overall `DCSexp` should match the corresponding slots of the component `DCSchr`s. 

```{r D03, include=T}
# Structure of a DCS experiment
str(Exp);
```

#### Methods

* A `summary` method is available for objects of class `DCSexp`. Two matrices are printed and returned. The first describes the pixels and maximal pixel separation for each chromosome in the experiment. The second describes the samples and their group assignments. 
* A `subset` method is available for objects of class `DCSchr`. Subsetting by either pixel or maximal pixel separation is available. To subset by pixel, supply a logical vector with length equal to the number of pixels. To subset by pixel separation, provided a (numeric) maximal separation. 
* A `subset` method is available for objects of class `DCSexp`. Subsetting by chromosome and maximal pixel separation is available. To subset by chromosome, specify the chromosomes to retain. To subset by pixel separation, provided a (numeric) maximal separation. 
* An `as.data.frame` method is available for objects of class `DCSchr` and `DCSexp`. 

```{r}
# Summary of DCS Experiment
cat("DCS Experiment Summary:\n");
S = summary(Exp);
cat("\n");
# Subsetting a DCS Chromosome using a logical vector
S = subset(x=A.1,pix=(A.1@Coord[,"i"]==80));
# Subsetting a DCS Chromosome
S = subset(x=A.1,dmax=10);
# Subsetting a DCS Experiment
S = subset(x=Exp,chrs=c("1"),dmax=10);
# Coercion to data.frame
cat("Head of Experiment Matrix, Chromosome 1, Max Separation 10, following Coercion:\n");
S.df = as.data.frame(S);
head(S.df);
```

## Focal Stat Objects

#### Structure

An object of class `FSchr` has the following slots:

* `@chr`, a string naming the chromosome.
* `@foci`, a numeric vector of foci. A **focus** refers to a pixel for which $i = j$, i.e. a pixel on the diagonal of the contact matrix. 
* `@groups`, a factor naming the groups. Data in an `FSchr` are summarized at the group level.
* `@Stats`, a list of data.frames with names corresponding to `@groups`. Each data frame is $n_{f}$ by $n_{q}$, where $n_{f}$ is the number of foci, and $n_{q}$ is the number of available statistics. 

```{r}
# Extracting chromosome one
B.1 = DCS::B@Data[["1"]];
# Structure of a FS chromosome
str(B.1);
```

An object of class `FSexp` has the following slots:

* `@chrs`, a character vector of chromosome names.
* `@Data`, a list of `FSchr`s, with names corresponding to `@chrs`.
* `@groups`, a factor naming the groups. Data in an `FSexp` are summarized at the group level.

Note that the `@groups` slots for the overall `FSexp` should match the corresponding slot of the component `FSchr`s.  

```{r}
# Structure of a FS experiment
str(DCS::B);
```

#### Methods

* A `subset` method is available for objects of class `FSchr`. Subsetting by focus and group is available. To subset by focus, supply a logical vector with length equal to the number of foci. To subset by group, specify the group(s) to retain.
* A `subset` method is available for objects of class `FSexp`. Subsetting by chromosome and group is available. To subset chromosome, specify the chromosome(s) to retain. To subset by group, specify the group(s) to retain.
* An `as.data.frame` method is available for objects of class `FSchr` and `FSexp`. 

```{r}
cat("Number of Foci:\n");
(nf = length(B.1@foci));
cat("\n");
# Subsetting a FSchr using a logical vector
S = subset(x=B.1,foci=c(rep(T,10),rep(F,nf-10)));
# Subsetting a FSchr by group
S = subset(x=B.1,group=c("0"));
# Subsetting a FSexp by chromsome and group
S = subset(x=B,chrs=c("1"),group=c("1"));
# Coercion to data.frame
cat("Head of Border Score Matrix, Chromosome 1, Group 1, following Coercion:\n");
S.df = as.data.frame(S);
head(S.df);
```
